---
title: "Socratic Tutor — where the class is struggling"
format:
  html:
    toc: true
    embed-resources: true
    fig-width: 8
    fig-height: 4.5
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup

library(DBI)
library(RPostgres)
library(dplyr)
library(ggplot2)
library(knitr)

# libpq accepts a full connection URI in `dbname`, so the Railway public
# connection string can be passed through unchanged. Set it before rendering:
#   Sys.setenv(TUTOR_DATABASE_URL = "postgresql://...")
# or put it in ~/.Renviron.
db_url <- Sys.getenv("TUTOR_DATABASE_URL")
if (!nzchar(db_url)) {
  stop("TUTOR_DATABASE_URL is not set. Copy the public connection string from ",
       "the Railway Postgres service and put it in ~/.Renviron.")
}

con <- dbConnect(RPostgres::Postgres(), dbname = db_url)

events <- dbGetQuery(con, "
  select e.student, e.task_id, e.level, e.question, e.found, e.ts,
         t.title, t.notebook
  from events e
  left join tasks t using (task_id)
  order by e.ts desc
")

theme_set(theme_minimal(base_size = 12))
```

```{r}
#| label: guard
#| results: asis

if (nrow(events) == 0) {
  cat("No tutor requests recorded yet. Once students start asking, this report fills in.\n\n")
  knitr::knit_exit()
}
```

Data through `r format(max(events$ts), "%Y-%m-%d %H:%M")`:
**`r nrow(events)`** requests from **`r dplyr::n_distinct(events$student)`** students
across **`r dplyr::n_distinct(events$task_id)`** tasks.

## Which tasks generate the most requests

The blunt measure of difficulty. Tall bars are tasks students cannot get through alone.

```{r}
#| label: requests-per-task

per_task <- events |>
  filter(found) |>
  count(task_id, title, sort = TRUE)

per_task |>
  head(20) |>
  mutate(label = paste0(task_id, if_else(is.na(title), "", paste0(" — ", title)))) |>
  ggplot(aes(x = n, y = reorder(label, n))) +
  geom_col(fill = "steelblue4") +
  labs(x = "requests", y = NULL)
```

## Which tasks leave students genuinely stuck

The more useful measure. A task with many requests but mostly level 1 is generating
quick clarifications. A task where students reach **level 3 or more** is one where
they are stuck — level is persistent, so it counts asks across separate chat
sessions, not just within one conversation.

```{r}
#| label: stuck-rate

stuck <- events |>
  filter(found) |>
  group_by(task_id, title) |>
  summarise(
    students   = n_distinct(student),
    reached_l3 = n_distinct(student[level >= 3]),
    .groups    = "drop"
  ) |>
  mutate(share = reached_l3 / students) |>
  filter(students >= 3) |>
  arrange(desc(share))

stuck |>
  head(20) |>
  mutate(label = paste0(task_id, if_else(is.na(title), "", paste0(" — ", title)))) |>
  ggplot(aes(x = share, y = reorder(label, share))) +
  geom_col(fill = "firebrick4") +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "share of students reaching level 3+", y = NULL,
       caption = "tasks with at least 3 students")
```

```{r}
#| label: stuck-table

stuck |>
  head(10) |>
  transmute(task_id, title, students, `reached L3+` = reached_l3,
            share = scales::percent(share, accuracy = 1)) |>
  kable()
```

## Who is asking

Four students asking sixty times and thirty students asking eight times are
different problems. The first is an outreach question, the second is a lab-design
question.

```{r}
#| label: concentration

events |>
  filter(found) |>
  count(student, sort = TRUE) |>
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 2, fill = "grey30") +
  labs(x = "requests per student", y = "students")
```

```{r}
#| label: top-students

events |>
  filter(found) |>
  group_by(student) |>
  summarise(requests = n(), tasks = n_distinct(task_id), last_seen = max(ts),
            .groups = "drop") |>
  arrange(desc(requests)) |>
  head(10) |>
  kable()
```

## When the work happens

```{r}
#| label: timing

events |>
  mutate(day = as.Date(ts)) |>
  count(day) |>
  ggplot(aes(x = day, y = n)) +
  geom_col(fill = "steelblue4") +
  labs(x = NULL, y = "requests")
```

## What students actually asked

Usually the most informative panel in this report: the wording of a confused
question tells you more than any count. Note that questions are recorded verbatim —
see the disclosure in the project README.

```{r}
#| label: recent-questions

events |>
  filter(!is.na(question), nzchar(question)) |>
  head(40) |>
  transmute(
    when = format(ts, "%m-%d %H:%M"),
    student, task_id, level,
    question = substr(question, 1, 220)
  ) |>
  kable()
```

## Tasks nobody asks about

Ambiguous by design: either everyone got it, or nobody attempted it. Worth pairing
with submission data before drawing a conclusion.

```{r}
#| label: silent-tasks

tasks <- dbGetQuery(con, "select task_id, title from tasks order by task_id")

anti_join(tasks, per_task, by = "task_id") |>
  kable()
```

```{r}
#| label: teardown

dbDisconnect(con)
```
