---
title: "R Lab: A template notebook for the Socratic Tutor"
format: html
author: Your Name
date: last-modified
execute:
  echo: true
toc: true
---

```{r}
#| label: load-packages
#| include: false
#| task: r-lab

# Any setup the student needs goes here.
# The `#| task:` directive at the top tells the tutor which notebook this is.
```

# Introduction

This file is a **template** for instructors adopting the Socratic Tutor. It shows
the structure the extension expects:

- A `#| task: <notebook-id>` marker at the top of the file (here, `r-lab`).
- One section per exercise, with a heading that contains the task ID in
  `{curly-braces}` (e.g. `{r-lab-1}`).
- A code chunk for the student, also tagged with `#| task: <exercise-id>`.
- A Quarto callout titled `"Solution"` containing the reference solution. The
  tutor reads this callout but never reveals it verbatim to the student.

Copy this file into your `notebook-solutions/` folder, rename it to match your
course (e.g. `r-lab.qmd`), and replace the two example exercises below with
your own.

# Sum the even numbers in a vector `{r-lab-1}`

**To-do:** Write a function `sum_even(x)` that returns the sum of all even
numbers in a numeric vector `x`. For example, `sum_even(1:6)` should return `12`.

A few hints:

- The modulo operator in R is `%%`.
- You can index a vector with a logical condition.

```{r}
#| label: sum-even
#| task: r-lab-1

# Your code here

```

::: {.callout-caution collapse="true" title="Solution"}

```{r}
#| label: sum-even-solution
#| task: r-lab-1
#| eval: false

sum_even <- function(x) {
  sum(x[x %% 2 == 0])
}

sum_even(1:6)  # 12
```

**Key points:**

- `x %% 2 == 0` produces a logical vector marking even entries.
- `x[<logical>]` keeps only the entries where the condition is `TRUE`.
- `sum()` then adds them up.
:::

# Count words in a string `{r-lab-2}`

**To-do:** Write a function `count_words(s)` that returns the number of words in
a character string `s`. Treat any run of whitespace as a single separator. For
example, `count_words("hello  world")` should return `2`.

```{r}
#| label: count-words
#| task: r-lab-2

# Your code here

```

::: {.callout-caution collapse="true" title="Solution"}

```{r}
#| label: count-words-solution
#| task: r-lab-2
#| eval: false

count_words <- function(s) {
  length(strsplit(trimws(s), "\\s+")[[1]])
}

count_words("hello  world")  # 2
```

**Key points:**

- `trimws()` removes leading/trailing whitespace so they don't produce empty tokens.
- `strsplit(s, "\\s+")` splits on one-or-more whitespace characters.
- `strsplit()` returns a list; `[[1]]` extracts the first (and only) element.
:::
